---
title: "Beaver Temperature"
author: Lea Keller
date: "`r date()`"
format:
  html:
    toc: true
    code-fold: show
    code-tools: true
    df-print: paged
    embed-resources: true
  pdf:
    toc: true
  docx:
    toc: true
---

# Set up

Loading all necessary packages and libraries.

```{r}
#| output: false
library(labelled)
library(tidyverse)
library(readr)
library(usethis)
library(here)
library(gtsummary)
library(dplyr)
library(lubridate)
library(ggplot2)
```

## Loading Data
Loading two different datasets with temperature data from two different beavers. The data is already included in R, so no need to import it from an external source.
```{r}
#Body Temperature Series of a beaver
beaver1 <- beaver1
beaver2 <- beaver2
```

## Data Wrangling
Data converted from military time to total minutes in order to create a continuous time variable for both days. The data was also cleaned and modified, because measurements were done after midnight, 1140 minutes were added to day 2 total minutes.
```{r}
#Calculate total minutes for each time point
beavers1_time <- beaver1 %>%
  mutate(
    hours = time %/% 100,
    minutes = time %% 100,
    total_minutes = (hours * 60) + minutes,
  )


beavers2_time <- beaver2 %>%
  mutate(
    hours = time %/% 100,
    minutes = time %% 100,
    total_minutes = (hours * 60) + minutes,
  )

#Clean and modify data, because measurments were done after midnight, 1140 minutes were added to day 2 total minutes. 
beavers1_time <- beavers1_time |>
  mutate(day = ifelse(day == 346,1,2)) |>
  mutate(total_minutes = ifelse(day== 2,
                       total_minutes + 1440,
                       total_minutes))
       

beavers2_time <- beavers2_time |>
  mutate(day = ifelse(day == 307,1,2)) |>
  mutate(total_minutes = ifelse(day== 2,
                                total_minutes + 1440,
                                total_minutes))
```

The two datasets were combined into one dataset for further analysis and plotting.
```{r}
#combine data
beavers_complete <- bind_rows(
  beavers1_time = beavers1_time,
  beavers2_time = beavers2_time,
.id = "beaver_id"
)
```

# Data Plotting
@fig-1 :Temperature plot beaver 1
```{r}
#| label: fig-1
#| fig-cap: "Temperature plot beaver 1"
#| fig-width: 8
#| fig-height: 4
plot_beavers1 <- ggplot(data = beavers1_time, 
                        mapping = aes(x = total_minutes, y = temp)) + 
  geom_point()
plot_beavers1
```
@fig-2: Temperature plot beaver 1
```{r}
#| label: fig-2
#| fig-cap: "Temperature plot beaver 2"
#| fig-width: 8
#| fig-height: 4
ggplot(data = beavers2_time, 
                        mapping = aes(x = total_minutes, y = temp)) + 
  geom_point()
```

@fig-3: Temperature of both beavers combined in a simple graph
```{r}
#| label: fig-3
#| fig-cap: "Combined temperature plot of both beavers"
#| fig-width: 8
#| fig-height: 4
ggplot(data = beavers_complete, 
                        mapping = aes(x = total_minutes, y = temp, color = beaver_id)) + 
  geom_line(mapping = aes(group = beaver_id ))
```
@fig-4: Temperature of both beavers combined in a more detailed graph with labelling and custom colours.
```{r}
#| label: fig-4
#| fig-cap: "Combined temperature plot of both beavers - with labelling"
#| fig-width: 8
#| fig-height: 4
beaver_plot <- ggplot(data = beavers_complete, 
                        mapping = aes(x = total_minutes, y = temp, colour = beaver_id)) + 
  geom_line(mapping = aes(group = beaver_id), 
            alpha = 0.7, linetype = "solid", linewidth = 1.5) +
  scale_colour_manual(name = "Beaver ID",
                      breaks = c("beavers1_time", "beavers2_time"), 
                      values = c("salmon", "purple"),
                      labels = c("Beaver 1", "Beaver 2")) + 
  ggtitle(label = "Body Temperature of two Beavers") +
  xlab(label = "Time (seconds)") +
  ylab(label = "Temperature (°C)")
beaver_plot 
```

# Save last plit as png file in output/figures folder.

```{r}
ggsave(beaver_plot, 
       filename = here("output", "figures", "beaver_plot.png"), 
       width = 8, height = 4, dpi = 300)
```


```{r}
sessionInfo()
```